---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(tree)
library(stats)
library(factoextra)
library(neuralnet)
library(mltools)
library(data.table)
library(MASS)
library(ISLR)
library(class)
library(caret)
```

```{r, include=FALSE}
cas <- read_csv("Cas.csv") %>%
  janitor::clean_names()
accidents <- read_csv("dftRoadSafety_Accidents_2016.csv") %>%
  janitor::clean_names()
makemodel <- read_csv("MakeModel2016.csv") %>%
  janitor::clean_names()
veh <- read_csv("Veh.csv") %>%
  janitor::clean_names()
```

# Introduction\
Here we need a summary of what we are trying to discover in our report

## The Dataset\
This should provide an analysis of the dataset and explore trends/patters in the data

**EDA**
```{r}
table(accidents$accident_severity)
```

## Questions of Interest\
After exploring the data, we decided to analyze two questions of interest.

1. Can We Help Find Safer Routes For Cyclists?
2. Can we predict when an accident will be fatal based on the driving conditions?


# Methodology\
This section is for describing the methodology we will use to analyze our questions of interest


## Data cleaning
```{r}
accidents <- accidents %>%
  mutate(weather_conditions = case_when(weather_conditions == -1 ~ "No data",
                                        weather_conditions == 1 ~ "Fine",
                                        weather_conditions == 2 ~ "Rain_no_high_wind",
                                        weather_conditions == 3 ~ "Snow_no_high_wind",
                                        weather_conditions == 4 ~ "Fine_high_wind",
                                        weather_conditions == 5 ~ "Rain_high_wind",
                                        weather_conditions == 6 ~ "Snow_high_wind",
                                        weather_conditions == 7 ~ "Fog_or_mist",
                                        weather_conditions == 8 ~ "Other",
                                        weather_conditions == 9 ~ "Unknown",
                                        TRUE ~ "NA"),
         accident_severity = case_when(accident_severity == 1 ~ "Fatal",
                                       accident_severity == 2 ~ "Serious",
                                       accident_severity == 3 ~ "Slight"),
         light_conditions = case_when(light_conditions == 1 ~ "Daylight",
                                      light_conditions == 4 ~ "Darkness_lights_lit",
                                      light_conditions == 5 ~ "Darkness_lights_unlit",
                                      light_conditions == 6 ~ "Darkness_no_lighting",
                                      light_conditions == 7 ~ "Unknown",
                                      light_conditions == -1 ~ "No data",
                                      TRUE ~ "NA"),
         road_surface_conditions = case_when(road_surface_conditions == 1 ~ "Dry",
                                      road_surface_conditions == 2 ~ "Wet_or_Damp",
                                      road_surface_conditions == 3 ~ "Snow",
                                      road_surface_conditions == 4 ~ "Frost_or_Ice",
                                      road_surface_conditions == 5 ~ "Flood_over_3cm",
                                      road_surface_conditions == 6 ~ "Oil_or_diesel",
                                      road_surface_conditions == 7 ~ "Mud",
                                      road_surface_conditions == -1 ~ "No data",
                                      TRUE ~ "NA")) %>%
  filter(weather_conditions != "No data" & weather_conditions != "Other" &
         weather_conditions != "Unknown" & weather_conditions != "NA" &
         light_conditions != "No data" & light_conditions != "Other" &
         light_conditions != "Unknown" & light_conditions != "NA" &
         road_surface_conditions != "No data" & road_surface_conditions != "Other" &
         road_surface_conditions != "Unknown" & road_surface_conditions != "NA") # get rid of ambiguous/unknown weather/light conditions
```



```{r}
    accidents1 <- accidents %>% sample_n(5000)
    str(accidents1)
    par(mfrow = c(2, 2))

    #accident_count <- table(accidents$accident_severity)
    #accident_count[1] / nrow(accidents$accident_severity)


    accidents1 %>%
        ggplot(aes(x = "", y = accident_severity, fill = accident_severity)) +
        geom_bar(stat = "identity", width = 1) +
        coord_polar(theta = "y", start = 0) +
        labs(x = "", y = "", title = "Proportion of accident severity")

    accidents1 %>%
        ggplot(aes(x = road_surface_conditions)) +
        geom_bar(aes(fill = accident_severity))

    accidents1 %>%
        ggplot(aes(x = speed_limit)) +
        geom_bar(aes(fill = accident_severity), position = "fill")

    accidents1 %>%
        ggplot(aes(x = urban_or_rural_area)) +
        geom_bar(aes(fill = accident_severity), position = "fill")



    accidents1 %>%
        ggplot(aes(x = light_conditions)) +
        geom_bar(aes(fill = accident_severity))

    accidents1 %>%
        ggplot(aes(x = weather_conditions)) +
        geom_bar(aes(fill = accident_severity), position = "fill")
```
